\LARGE \bf Preparation of Papers for IEEE Sponsored Conferences \& Symposia*

### 2017-04-08 bmc.tex
.

\inclfig[width=0.95\linewidth]{vdp_disc_map.pdf}{Cells and
trajectory segments used by $1$-relational modeling.}
\label{fig:vdp-map}

@@

\begin{table}[!htbp]
\centering
\caption{PWA model computed using OLS. The affine model for each edge $(C,C')$ in the graph \figref{vdp-graph} is given by $x'\in Ax + b + \delta$, where $\delta$ is a vector of intervals.}
\label{tab:vdp-pwa}
\begin{tabular}{@{}ccccc@{}}
\toprule
$C$ & $C'$ & $A$ & $b$ & $\delta$\\
\midrule
(0, -1)   & (0, -1)   & $\begin{bmatrix}0.99& 0.12 \\-0.12&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
(0, -1)   & (0, -2)   & $\begin{bmatrix}0.99& 0.12 \\-0.10&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
\midrule
(0, -2)   & (0, -2)   & $\begin{bmatrix}0.99& 0.12 \\-0.09&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
(0, -2)   & (0, -4)   & $\begin{bmatrix}0.99& 0.12 \\-0.07&1.62\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
\midrule
(0, -9)   & (-1, -15) & $\begin{bmatrix}0.99& 0.12 \\-0.09&1.61\end{bmatrix}$&$\begin{bmatrix}0\\-0.04\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
\midrule
(-1, -15) & (-3, -24) & $\begin{bmatrix}0.95& 0.12 \\-1.29&1.47\end{bmatrix}$&$\begin{bmatrix}-0.01\\-0.41\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
\midrule
(-3, -24) & (-4, -29) & $\begin{bmatrix}1.11&-0.17 \\-1.71&0.66\end{bmatrix}$&$\begin{bmatrix}-1.07\\-3.31\end{bmatrix}$&$\begin{bmatrix}[-0.04, 0.04]\\ [-0.07, 0.09]\end{bmatrix}$\\
\midrule
(-4, -29) & (-4, -29) & $\begin{bmatrix}0.99& 0.01 \\-0.41&1.02\end{bmatrix}$&$\begin{bmatrix}0\\-0.28\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
\midrule
(-1, -10) & (-2, -16) & $\begin{bmatrix}0.97& 0.12 \\-0.71&1.56\end{bmatrix}$&$\begin{bmatrix}0\\-0.11\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
(-1, -10) & (-2, -15) & $\begin{bmatrix}0.97& 0.12 \\-0.70&1.58\end{bmatrix}$&$\begin{bmatrix}0\\-0.08\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
\midrule
(-2, -16) & (-3, -23) & $\begin{bmatrix}0.92& 0.12 \\-1.84&1.39\end{bmatrix}$&$\begin{bmatrix}-0.02\\-0.73\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
\midrule
(-3, -23) & (-5, -29) & $\begin{bmatrix}3.14&-0.81 \\-0.96&0.23\end{bmatrix}$&$\begin{bmatrix}-3.13\\-4.99\end{bmatrix}$&$\begin{bmatrix}[-0.04, 0.03]\\ [-0.03, 0.04]\end{bmatrix}$\\
\midrule
(-2, -15) & (-3, -22) & $\begin{bmatrix}0.93& 0.12 \\-1.75&1.40\end{bmatrix}$&$\begin{bmatrix}-0.02\\-0.68\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
\midrule
(-3, -22) & (-5, -29) & $\begin{bmatrix}3.08&-0.72 \\-1.69&0.40\end{bmatrix}$&$\begin{bmatrix}-2.77\\-4.57\end{bmatrix}$&$\begin{bmatrix}[-0.02, 0.02]\\ [-0.01, 0.02]\end{bmatrix}$\\
\bottomrule
\end{tabular}
\end{table}

@@

The search parameters for S3CAM-R include the parameters of S3CAM:
$N$, $\epsilon$ and $\Delta$. Additionally, they also include the
maximum error budget $\delta_{max}$ for OLS and the maximum length of
segmented trajectory for building $k$-relational models.

We have already discussed the effects of $N$, $\epsilon$ and $\Delta$
on S3CAM's performance. However, they also have an effect on
relational modeling. A finer grid with small cells produces more
accurate models than a coarser grid with large cells. Similarly,
small time length trajectory segments result in more accurate
modeling.
%An increase in the number of samples can increases the
%accuracy of regression.

>>>>

\inclfig[width=0.90\linewidth]{vdp_disc_map.pdf}{Cells and
trajectory segments used by $1$-relational modeling.}
\label{fig:vdp-map}

@@

\begin{table}[h]
\centering
\caption{PWA model computed using OLS. The affine model for each edge $(C,C')$ in the graph \figref{vdp-graph} is given by $x'\in Ax + b + \delta$, where $\delta$ is a vector of intervals.}
\label{tab:vdp-pwa}
\begin{tabular}{@{}ccccc@{}}
\toprule
$C$ & $C'$ & $A$ & $b$ & $\delta$\\
\midrule
(0, -1)   & (0, -1)   & $\begin{bmatrix}0.99& 0.12 \\-0.12&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
(0, -1)   & (0, -2)   & $\begin{bmatrix}0.99& 0.12 \\-0.10&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
\midrule
(0, -2)   & (0, -2)   & $\begin{bmatrix}0.99& 0.12 \\-0.09&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
(0, -2)   & (0, -4)   & $\begin{bmatrix}0.99& 0.12 \\-0.07&1.62\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
\midrule
(0, -9)   & (-1, -15) & $\begin{bmatrix}0.99& 0.12 \\-0.09&1.61\end{bmatrix}$&$\begin{bmatrix}0\\-0.04\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
\midrule
(-1, -15) & (-3, -24) & $\begin{bmatrix}0.95& 0.12 \\-1.29&1.47\end{bmatrix}$&$\begin{bmatrix}-0.01\\-0.41\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
\midrule
(-3, -24) & (-4, -29) & $\begin{bmatrix}1.11&-0.17 \\-1.71&0.66\end{bmatrix}$&$\begin{bmatrix}-1.07\\-3.31\end{bmatrix}$&$\begin{bmatrix}[-0.04, 0.04]\\ [-0.07, 0.09]\end{bmatrix}$\\
\midrule
(-4, -29) & (-4, -29) & $\begin{bmatrix}0.99& 0.01 \\-0.41&1.02\end{bmatrix}$&$\begin{bmatrix}0\\-0.28\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
\midrule
(-1, -10) & (-2, -16) & $\begin{bmatrix}0.97& 0.12 \\-0.71&1.56\end{bmatrix}$&$\begin{bmatrix}0\\-0.11\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
(-1, -10) & (-2, -15) & $\begin{bmatrix}0.97& 0.12 \\-0.70&1.58\end{bmatrix}$&$\begin{bmatrix}0\\-0.08\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
\midrule
(-2, -16) & (-3, -23) & $\begin{bmatrix}0.92& 0.12 \\-1.84&1.39\end{bmatrix}$&$\begin{bmatrix}-0.02\\-0.73\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
\midrule
(-3, -23) & (-5, -29) & $\begin{bmatrix}3.14&-0.81 \\-0.96&0.23\end{bmatrix}$&$\begin{bmatrix}-3.13\\-4.99\end{bmatrix}$&$\begin{bmatrix}[-0.04, 0.03]\\ [-0.03, 0.04]\end{bmatrix}$\\
\midrule
(-2, -15) & (-3, -22) & $\begin{bmatrix}0.93& 0.12 \\-1.75&1.40\end{bmatrix}$&$\begin{bmatrix}-0.02\\-0.68\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
\midrule
(-3, -22) & (-5, -29) & $\begin{bmatrix}3.08&-0.72 \\-1.69&0.40\end{bmatrix}$&$\begin{bmatrix}-2.77\\-4.57\end{bmatrix}$&$\begin{bmatrix}[-0.02, 0.02]\\ [-0.01, 0.02]\end{bmatrix}$\\
\bottomrule
\end{tabular}
\end{table}

@@

The search parameters for PWA-Rel include the parameters of S3CAM:
$N$, $\epsilon$ and $\Delta$. Additionally, they also include the
maximum error budget $\delta_{max}$ for OLS and the maximum length of
segmented trajectory for building $k$-relational models.

The effects of $N$, $\epsilon$ and $\Delta$
on S3CAM's performance were detailed in~\cite{zutshi2014multiple}. However, they also have an effect on
relational modeling. A finer grid with small cells produces more
accurate models than a coarser grid with large cells. Similarly,
small time length trajectory segments result in more accurate
modeling.
%An increase in the number of samples can increases the
%accuracy of regression.

### 2017-04-08 implementation.tex
.

The implementation was prototyped as S3CAM-R, an extension to our
previously mentioned tool S3CAM \cite{zutshi2014multiple}. OLS regression
routines were used from Scikit-learn~\cite{pedregosa2011scikit}, a
Python module for machine learning. SAL~\cite{SAL-SRI}
with Yices2~\cite{dutertre2014yices} was the model checker and the SMT
solver.

@@

We tabulate our preliminary evaluation in \tabref{res-rel}. We used
few of the benchmarks described in \cite{zutshi2014multiple}, including the Van
der Pol oscillator, Brusselator, Lorenz attractor and the Navigation
benchmark. As before, we ran S3CAM-R $10$ times with different seeds
and averaged the results. We tabulate both the total time taken and
the time taken by SAL to compute the counter-example and compare
against a newer implementation of S3CAM. Note that different
abstraction parameters are used for S3CAM and S3CAM-R, due to the
differences in which they operate.

\begin{table}[!htbp]
\centering
\caption{Avg. timings for benchmarks. The \textbf{BMC} column lists
time taken by the BMC engine. The total time in seconds (rounded off
to an integer) is noted under \textbf{S3CAM-R} and \textbf{S3CAM}.
$TO$ signifies time $>5hr$, after which the search was killed.}
\label{tab:res-rel}
\begin{tabular}{@{}llll@{}}
\toprule
    Benchmark & PWA Rel + BMC & PWA Rel + LP & S3CAM\\
\midrule
    Van der Pol ($\prop_3$)   &$14$ & $20$ & $24$\\
    Brusselator               &$4$ & $4$ & $2$\\
    Lorenz                    &$122$ & $120$ & $35$\\
%B.Ball                    &  &   & \\
    Nav ($\prop_P$)           &$TO$ & $1710$  & $603$ \\%9423 + 4480
    Nav ($\prop_Q$)           &$TO$ & $1560$   & $546$ \\%8105 + 970
    Nav ($\prop_R$)           &$TO$ & $2352$  & $2003$\\
    Nav ($\prop_S$)           &$TO$ & $0$  & $2100$\\
    Bouncing Ball             &$Safe$ & $Safe$  & $450$\\

@@

The results show promise, but clearly S3CAM performs better. Also,
S3CAM-R timed out on some benchmarks like Nav $\prop_R$, Nav $\prop_S$
and the bouncing ball. However, we need to explore more benchmarks to
be conclusive. S3CAM-R's performance can be explained by the
difficulty in finding a good abstraction ($\quant_\epsilon$), which
needs to be fine enough, to obtain good prediction models but coarse
enough, to be manageable by the current SMT solvers. Specifically, to
obtain a good quality counter-example from the model checker (and
avoid false positives), the transition system should be created from
models with high accuracy, for which a finer abstraction is required.
However, the exploration of a finer abstraction is exponentially more
complex and results in a much bigger transition system, which the
current state of the art bounded model checkers (which use SMT
solvers) can not handle under reasonable resources.

>>>>

The implementation was prototyped as PWA-Rel,
%an extension to our
%previously mentioned tool S3CAM \cite{zutshi2014multiple}.
OLS regression
routines were used from Scikit-learn~\cite{pedregosa2011scikit}, a
Python module for machine learning. SAL~\cite{SAL-SRI}
with Yices2~\cite{dutertre2014yices} was the model checker and the SMT
solver.

@@

We tabulate our preliminary evaluation in \tabref{res-rel}. We used
few of the benchmarks described in \cite{zutshi2014multiple},
including the Van der Pol oscillator, Brusselator, Lorenz attractor
and the Navigation benchmark. As before, we ran PWA-Rel $10$ times
with different seeds and averaged the results. We tabulate both the
total time taken and the time taken by SAL to compute the
counter-example and compare against an implementation of S3CAM. Note
that different abstraction parameters are used for S3CAM and
PWA-Rel, due to the differences in which they operate.

\begin{table}[!htbp]
\centering
\caption{Avg. timings for benchmarks. The \textbf{BMC} column lists
time taken by the BMC engine. The total time in seconds (rounded off
to an integer) is noted under \textbf{PWA-Rel} and \textbf{S3CAM}.
$TO$ signifies time $>5hr$, after which the search was killed.}
\label{tab:res-rel}
\begin{tabular}{@{}llll@{}}
\toprule
    Benchmark & PWA Rel + BMC & PWA Rel + LP & S3CAM\\
\midrule
    Van der Pol ($\prop_3$)   &$14$ & $20$ & $24$\\
    Brusselator               &$4$ & $4$ & $2$\\
    Lorenz                    &$122$ & $120$ & $35$\\
%B.Ball                    &  &   & \\
    Nav ($\prop_P$)           &$TO$ & $1710$  & $603$ \\%9423 + 4480
    Nav ($\prop_Q$)           &$TO$ & $1560$   & $546$ \\%8105 + 970
    Nav ($\prop_R$)           &$TO$ & $2352$  & $2003$\\
    Nav ($\prop_S$)           &$TO$ & $0$  & $2100$\\
    Bouncing Ball             &$Safe$ & $Safe$  & $450$\\

@@

The results show promise, but the performance is comparable. Also,
PWA-Rel timed out on some benchmarks like Nav $\prop_R$, Nav $\prop_S$
and the bouncing ball. However, we need to explore more benchmarks to
be conclusive. PWA-Rel's performance can be explained by the
difficulty in finding a good abstraction ($\quant_\epsilon$), which
needs to be fine enough, to obtain good prediction models but coarse
enough, to be manageable by the current SMT solvers. Specifically, to
obtain a good quality counter-example from the model checker (and
avoid false positives), the transition system should be created from
models with high accuracy, for which a finer abstraction is required.
However, the exploration of a finer abstraction is exponentially more
complex and results in a much bigger transition system, which the
current state of the art bounded model checkers (which use SMT
solvers) can not handle under reasonable resources.
